Towards a Cognitive System That Can Recognize Spatial Regions Based on Context

### 2012-01-25 scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex
Kate: Added Langly ref, Redid Analogy Diagram.  Minor tweaking of conclusion

The ability to understand and reason about \textit{spatial regions} is essential for cognitive systems performing tasks for humans in everyday environments. Some regions, such as whole rooms and corridors, are defined by clearly perceivable boundaries (e.g. walls and doors). However, many regions to which humans routinely refer are not so easily defined. Consider, for example, the aforementioned region \textit{the front of the classroom}. This region is not easily perceivable using just the geometry of the environment. Instead, it is defined by the objects present in the room (chairs, a desk, a whiteboard), their role in this context (seats for students to watch a teacher who writes on the whiteboard) and their configuration in space (the seats point toward the whiteboard). We refer to such regions as \textit{context-dependent spatial regions} (CDSRs).

Current cognitive systems are not capable of representing and reasoning about CDSRs, yet it is an important ability. If cognitive systems are to collaborate with humans in everyday environments then they must be able to understand the spatial regions humans refer to. Many regions are best defined in a context-dependent manner, for example, a kitchen in a studio apartment, an aisle in a church or store, behind enemy lines in a military engagement, etc. In order to represent and reason about such regions, cognitive systems must integrate different types of information, including geometric, semantic, and functional knowledge. Creating systems able to integrate such a range of information is a key challenge in the cognitive systems paradigm.

This paper presents an artificial cognitive system (specifically a mobile robot) able to represent and reason about CDSRs. Our approach is founded on two assumptions. The first assumption is that CDSRs can be defined using \emph{qualitative spatial representations} (QSRs) extracted from the sensor data of the system~\cite{Cohn:2001}. The second assumption is that semantically similar areas (e.g. two different classrooms) will feature similar CDSRs, and that these similarities can be recognised through \emph{analogy}. The rest of the paper is structured following these assumptions. Section~\ref{sec:qsr-gen} describes how we generate QSRs from sensor data taken from an existing, state-of-the-art, cognitive system and use these to define CDSRs. Section~\ref{sec:analogy} then describes how we use the structure-mapping model of analogy \cite{Gentner1983a} to transfer a CDSR from a labelled example to a new situation. Section~\ref{sec:example} presents a worked example of the entire process, and Section~\ref{sec:evaluation} evaluates our system in comparison to data from human subjects performing the same task.

@@

Typical approaches to spatial representation for mobile robots tend to focus on localization, and thus mostly represent the world uniformly without subdivision into meaningful (semantic) units \cite{Thrun02a}. When a more structured representation is required, many turn to Kuipers' Spatial Semantic Hierarchy~\cite{Kuipers:2000}. This paper follows in this tradition, adding CDSRs to his qualitative topological representations. Whilst mobile robots exist which can determine the type of a room from the objects in it ~\cite{Hanheide/etal:2010a,Galindo/etal:2005a}, they only concern themselves with types of whole rooms, and cannot represent regions within rooms. This is also true for those systems which use some elements of QSR~\cite{aydemir2011icra}. The need for an autonomous system to ground references to human-generated descriptions of space has been recognised in domains where a robot must be instructed to perform a particular task, however existing systems are restricted to purely geometrically-defined regions~\cite{Tellex:2011,Dzifcak/etal:2009,brenneretal07ijcai}.

@@

In this paper we presented an integrated cognitive system capable of representing and reasoning about context-dependent spatial regions. The system is able to use analogy to recognise CDSRs in previously unseen environments through a process of analogy with a manually annotated known room. The system demonstrates a successful integration of a range of technologies including vision, SLAM, qualitative spatial reasoning and analogy. In order to make this rich collection of components work together, our work takes a number of short-cuts that we plan to address with in future work. These include a reliance on the initial orientation of a room in a global coordinate frame, the lack of a mechanism to retrieve relevant rooms from memory (e.g. MAC/FAC \cite{forbus/etal1995}), and a lack of transfer post-processing (e.g. comparing the QSRs present in both base and transferred regions) to improve results. Despite these limitations our evaluation demonstrated that the system is able to match the CDSRs produced by trial subjects on unseen rooms to an acceptable degree \textbf{SOMENUMBER}.

% \section{Acknowledgments}
% 
% The authors would like to thank Jeffery Usher for providing the code to compute Voronoi diagrams. The research leading to these results has received funding from the European Community's Seventh Framework Programme [FP7/2007-2013] under grant agreement No. 215181, CogX.

>>>>

The ability to understand and reason about \textit{spatial regions} is essential for cognitive systems performing tasks for humans in everyday environments. Some regions, such as whole rooms and corridors, are defined by clearly perceivable boundaries (e.g. walls and doors). However, many regions to which humans routinely refer are not so easily defined. Consider, for example, the aforementioned region \textit{the front of the classroom}. This region is not perceivable using just the geometry of the environment. Instead, it is defined by the objects present in the room (chairs, a desk, a whiteboard), their role in this context (seats for students to watch a teacher who writes on the whiteboard) and their configuration in space (the seats point toward the whiteboard). We refer to such regions as \textit{context-dependent spatial regions} (CDSRs).

Current cognitive systems are not capable of representing and reasoning about CDSRs, yet it is an important ability. If cognitive systems are to collaborate with humans in everyday environments then they must be able to understand and refer to the same spatial regions humans do. Many regions are best defined in a context-dependent manner, for example, a kitchen in a studio apartment, an aisle in a church or store, behind enemy lines in a military engagement, etc. In order to represent and reason about such regions, cognitive systems must integrate different types of information, including geometric, semantic, and functional knowledge. Creating systems able to integrate such a range of information is a key challenge in the cognitive systems paradigm \cite{langley:inpress}.

This paper presents an artificial cognitive system (specifically a mobile robot) able to represent and reason about CDSRs. Our approach is founded on two assumptions. The first assumption is that CDSRs can be defined using \emph{qualitative spatial representations} (QSRs) corresponding to sensor data of the system~\cite{Cohn:2001}. The second assumption is that semantically and geometrically similar areas (e.g. two different classrooms) will feature similar CDSRs, and that these similarities can be recognised through \emph{analogy}. The rest of the paper is structured following these assumptions. Section~\ref{sec:qsr-gen} describes how we generate QSRs from sensor data taken from an existing, state-of-the-art, cognitive system and use these to define CDSRs. Section~\ref{sec:analogy} then describes how we use the structure-mapping model of analogy \cite{Gentner1983a} to transfer a CDSR from a labelled example to a new situation. Section~\ref{sec:example} presents a worked example of the entire process, and Section~\ref{sec:evaluation} evaluates our system in comparison to data from human subjects performing the same task.

@@

Typical approaches to spatial representation for mobile robots tend to focus on localization, and thus mostly represent the world uniformly without subdivision into meaningful (semantic) units \cite{Thrun02a}. When a more structured representation is required, many turn to Kuipers' Spatial Semantic Hierarchy~\cite{Kuipers:2000}. This paper follows in this tradition, adding CDSRs to his qualitative topological representations. Whilst mobile robots exist which can determine the type of a room from the objects in it ~\cite{Hanheide/etal:2010a,Galindo/etal:2005a}, they only concern themselves with types of whole rooms, and cannot represent regions within rooms. This is also true for those systems which use some elements of QSR~\cite{aydemir2011icra}. The need for an autonomous system to ground references to human-generated descriptions of space has been recognized in domains where a robot must be instructed to perform a particular task, however existing systems are restricted to purely geometrically-defined regions~\cite{Tellex:2011,Dzifcak/etal:2009,brenneretal07ijcai}.

@@

In this paper we presented an integrated cognitive system capable of representing and reasoning about context-dependent spatial regions. The system identifies CDSRs in previously unseen environments through analogy with a manually annotated known room. The system demonstrates a successful integration of a range of technologies including vision, SLAM, qualitative spatial reasoning and analogy. In order to make this rich collection of components work together, our work takes a number of short-cuts that we plan to address with future work. These include a reliance on the initial orientation of a room in a global coordinate frame, the lack of a mechanism to retrieve relevant rooms from memory (e.g. MAC/FAC \cite{forbus/etal1995}), and a lack of transfer post-processing (e.g. comparing the QSRs present in both base and transferred regions) to improve results. Despite these limitations our evaluation demonstrated that the system is able to match the CDSRs produced by trial subjects on unseen rooms to an acceptable degree \textbf{SOMENUMBER}.

% \section{Acknowledgments}
% 
% The authors would like to thank Jeffery Usher for providing the code to compute Voronoi diagrams, Scott Friedman for assistance on probablistic SME, and Danny Bobrow comments on an earlier draft of this paper. The research leading to these results has received funding from the European Community's Seventh Framework Programme [FP7/2007-2013] under grant agreement No. 215181, CogX.
